### 2016-05-14 paper/paper.tex
Finished results section

\title{A Comparison of Dimensionality Reduction Techniques for Robotic Motion
Planning}

\date{\today}

\author{Brandon Araki \and Alex Wallar}

\documentclass[12pt]{article}

\usepackage[pdftex]{graphicx}
\graphicspath{{figs/}}
\usepackage{subcaption}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathabx}

\usepackage{mathptmx}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}

\usepackage{relsize}

\usepackage{float}

\usepackage{algorithm}

\usepackage[noend]{algorithmic}

\floatstyle{ruled} \newfloat{program}{thp}{lop} \floatname{program}{Structure}

\newcommand{\Normal}[3]{\mathcal{N}(#1, #2, #3)}

\newcommand{\Acronym}[1]{\ensuremath{{\small{\texttt{#1}}}}}
\newcommand{\Name}{\Acronym{Camgaze.js}} \newcommand{\False}{\Constant{false}}
\newcommand{\True}{\Constant{true}}
\newcommand{\Symbol}[1]{\ensuremath{\mathcal{#1}}}
\newcommand{\Function}[1]{\ensuremath{{\small \textsc{#1}}}}
\newcommand{\Constant}[1]{\ensuremath{\small{\texttt{#1}}}}
\newcommand{\Var}[1]{\ensuremath{{\small{\textsl{#1}}}}}
\newcommand{\argmin}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{min}}\;}
\newcommand{\grad}[1]{\underset{#1}{\operatorname{\Function{GradientDecent}}}\;}

\newcommand{\xr}[0]{\ensuremath{{x_{\Var{rand}}}}}
\newcommand{\xnear}[0]{\ensuremath{{x_{\Var{near}}}}}
\newcommand{\xnearest}[0]{\ensuremath{{x_{\Var{nearest}}}}}
\newcommand{\xnew}[0]{\ensuremath{{x_{\Var{new}}}}}
\newcommand{\xmin}[0]{\ensuremath{{x_{\Var{min}}}}}
\newcommand{\xparent}[0]{\ensuremath{{x_{\Var{parent}}}}}
\newcommand{\cmin}[0]{\ensuremath{{c_{\Var{min}}}}}
\newcommand{\cost}[0]{\ensuremath{{\Function{Cost}}}}
\newcommand{\linef}[1]{\ensuremath{{c(\Function{Line}(#1))}}}

\begin{document}

\maketitle

\section{Introduction}

Dimensionality reduction is a key component of making robotic motion planning
fast and efficient. A rigid body has six degrees of freedom (DoF), so a
multi-link robot such as a humanoid or snake robot can have dozens or hundreds
of DoF. The rotational and translational transformations of a rigid body can be
described with the 3-dimensional Special Euclidean group (known as SE(3)).
SE(3) is homeomorphic to the topological space
$\mathbb{R}^3\bigtimes\mathbb{RP}^3$ (where $\mathbb{RP}^3$ is the
3-dimensional real projective plane). Therefore it is easy to imagine a
multi-link robot \cite{kuindersma2015optimization} or a multi-robot system
\cite{alonso2015multi} with an arbitrarily complex state space. The picture is
further complicated by the addition of obstacles into a robot's world. The
state space for motion planning, which must take into account both the robot
and obstacles, is called the configuration space (or C-space)
\cite{lozano1983spatial}.

Motion planning essentially consists of a search over the configuration space
from a start configuration to a goal configuration. A huge number of methods
for searching through the configuration space have been developed, most of
which can be divided into two classes, sampling-based motion planning and
combinatorial motion planning \cite{lavalle2006planning}. In sampling-based
motion planning, obstacles in the C-space are defined implicitly so that the
path is constructed by randomly or pseudorandomly sampling points from the
C-space, doing collision checking, and connecting valid points to their nearest
neighbors. In combinatorial motion planning, obstacles are defined explicitly
and a complete search of the C-space is made. Because of the difficulty of
explicitly defining all of the geometry in a potentially complex world,
sampling-based motion planning techniques such as the Rapidly expanding Random
Tree (RRT) and the Probabilistic RoadMap (PRM) are the most popular motion
planning techniques.

We used PRM and RRTConnect in our project.

\subsection{PRM}

The Probabilistic RoadMap algorithm has two phases: the preprocessing phase in
which a roadmap is constructed by sampling random points in the C-space, and
the query phase in which a graph search over the roadmap is made to connect
initial and goal configurations \cite{kavraki1996prm}. The map is constructed
during the preprocessing phase by sampling random points in the C-space and
then adding them to the vertices in the roadmap within a certain radius of the
point. The algorithm is described in Alg~\ref{alg:PRM}.

\begin{algorithm}[h!] 
    \caption{$\Function{PRM}$~\cite{karaman2011sampling}}
    \label{alg:PRM}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}

        \STATE $V \leftarrow \{x_{\Var{init}}\} \cup \{\Var{SampleFree}_i\}_{i=1,\ldots,n}$
        \STATE $E \leftarrow \emptyset$
        \FORALL{$v \in V$}
            \STATE $U \leftarrow \Function{NearestNeighbors}(v)$
            \FORALL{$u \in U $}
                \IF{$\Function{CollisionFree}(v, u)$}
                    \STATE $E \leftarrow E \cup \{(v, u), (u, v)\}$
                \ENDIF
            \ENDFOR
        \ENDFOR
        \RETURN $G = (V, E)$

    \end{algorithmic}
\end{algorithm}

\subsection{RRTConnect}

\begin{algorithm}[h!] 
    \caption{$\Function{Extend\_RRT($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
    \label{alg:extend_RRT}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}
        \STATE $ q_{near} \leftarrow \Function{NEAREST\_NEIGHBOR($q$,$\;\mathcal{T}$)} $
        \IF{ $ \Function{NEW\_CONFIG($q$,$\;q_{near}$,$\;q_{new}$)} $}
            \STATE $ \mathcal{T}.\text{add\_vertex($q_{new}$)}; $
            \STATE $ \mathcal{T}.\text{add\_edge($q_{near}$,$\;q_{new}$)}; $
            \IF{$ q_{new} = q $}
                \RETURN $Reached$;
            \ELSE
                \RETURN $Advanced$;
            \ENDIF
        \ENDIF
        \RETURN $Trapped$;
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[h!] 
    \caption{$\Function{Connect\_RRT($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
    \label{alg:connect_RRT}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}
        \REPEAT 
            \STATE $S \leftarrow \Function{EXTEND\_RRT($\mathcal{T}$,$\;q$)};$ 
        \UNTIL {$ \textbf{not} \; (S = Advanced)$}
        \RETURN $S;$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[!h] 
    \caption{$\Function{RRTConnect($q_{init}$,$\;q_{goal}$)}$~\cite{kuffner2000rrt}}
    \label{alg:RRTConnect}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}
        \STATE $\mathcal{T}_a$.init($q_{init}$); $ \mathcal{T}_b$.init($q_{goal}$);
        \FOR{$k = 1 \ \TO \ K$}
            \STATE $q_{rand} \leftarrow \Function{RANDOM\_CONFIG()}$;
            \IF {$\textbf{not} \ (\Function{EXTEND\_RRT($\mathcal{T}_a$,$\;q_{rand} = Trapped$)})$}
                \IF {$ (\Function{CONNECT\_RRT($\mathcal{T}_b,q_{new}$}) = Reached $}
                    \RETURN $ \Function{PATH($\mathcal{T}_a,\mathcal{T}_b$)} $
                \ENDIF
            \STATE \Function{SWAP($\mathcal{T}_a,\mathcal{T}_b)$}
            \ENDIF
        \ENDFOR
        \RETURN $Failure$
    \end{algorithmic}
\end{algorithm}

The Rapidly exploring Random Tree algorithm works by building up a tree from
the initial configuration to the goal configuration \cite{lavalle1998rrt}.
Starting with the initial configuration as the root, it incrementally attempts
to add randomly selected samples to the tree until it reaches the goal
configuration.  The tree is ``rapidly exploring'' because the sample selection
is biased towards areas with large Voronoi regions (where the Voronoi regions
are computed based off of the nodes of the tree). RRTConnect, which is
described in Algos.~\ref{alg:extend_RRT},~\ref{alg:connect_RRT},
and~\ref{alg:RRTConnect}, differs from RRT in that two trees, one with its root
at the initial configuration and one with its root at the start configuration,
grow towards each other \cite{kuffner2000rrt}.

\section{Problem Description}

Despite the efficiency of sampling-based search, the high dimensionality of the
configuration space of complex robots means that motion planning for many
robots is slow or infeasible. This is because of the ``curse of dimensionality",
in which the size of the search space increases exponentially with the
dimension. One way to speed up motion planning is to reduce the dimensionality
of the C-space and do search in this reduced dimensionality space. For this
project, we tested the performance of a number of dimensionality reduction
techniques to see which one was the most effective in speeding up motion
planning while providing reliable obstacle avoidance.

\section{Dimensionality Reduction}

We used Feature Agglomeration, Truncated SVD, PCA, and Randomized PCA to
perform dimensionality reduction. We will briefly summarize each of these
techniques.

\subsection{PCA}

Principle Component Analysis, or PCA, transforms a set of variables into their
``principal components", or linearly uncorrelated variables
\cite{shlens2014pca}. Given an $m \times n$ matrix $\mathbf{X}$, you can find
its principal components by finding an orthonormal matrix $\mathbf{P}$ in $
\mathbf{Y} = \mathbf{PX} $ such that $\mathbf{C_Y} = \mathbf{YY}^T/n$. The rows
of $\mathbf{P}$ are the principal components of $\mathbf{X}$. In other words,
the axes of the dataset in question are reoriented so that the covariance of
the data along the first axis is minimized. After the first axis has been set,
the second axis is then set to minimize the covariance of the dataset along
that axis, and so on, so that the first axis has the lowest covariance, the
second axis has the second lowest, and so on. Dimensionality reduction is then
performed by chopping off the axes with the high covariance.

\subsection{Randomized PCA}

In Randomized PCA, computations are sped up by only approximating the singular
vectors of the components.

\subsection{Feature Agglomeration}

Feature Agglomeration is a form of hierarchical clustering in which clusters
are merged or split successively \cite{featureagglomeration}. Feature
Agglomeration groups together similar features to reduce the number of
features. It does this by using agglomerative clustering, in which every data
point starts as a cluster, and clusters are merged together.

\subsection{Truncated SVD}

In singular value decomposition, a matrix $\mathbf{M}$ is factorized into the
form $\mathbf{U} \Sigma \mathbf{V^*}$. $\Sigma$ is a diagonal matrix, and the
entries of $\Sigma$ are the singular values of $\mathbf{M}$. In Truncated SVD,
$\mathbf{M}$ is approximated by matrices $\mathbf{U_k} \Sigma_k
\mathbf{V^*_k}$, where $ k < r $, $r$ being the number of non-zero singular
values of the matrix \cite{maciejewski1989svd}. In effect you are approximating
$\mathbf{M}$ by cutting off low-value singular values. In this sense Truncated
SVD is similar to PCA; however, instead of cutting off the lowest eigenvalues
of the coviarance matrix, you are cutting off the singular values of the
matrix.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.32\textwidth]{2d_example_1.pdf}
    \includegraphics[width=0.32\textwidth]{2d_example_2.pdf}
    \includegraphics[width=0.32\textwidth]{2d_example_3.pdf} \\
    \includegraphics[width=0.32\textwidth]{3d_example_1.pdf}
    \includegraphics[width=0.32\textwidth]{3d_example_2.pdf}
    \includegraphics[width=0.32\textwidth]{3d_example_3.pdf}
    \caption{Examples of paths found in 2D and 3D scenes of random obstacles
    using the PRM algorithm. The green and red circles indicate the initial
    and goal configurations respectively. The blue path indicates the candidate
    solution through the space that connects the initial and goal
    configurations.}
    \label{fig:examples}
\end{figure}

\section{Methods}

\subsection{Simplifications}

Although our aim is to show how dimensionality reduction techniques can speed
up motion planning for robots, doing motion planning for an actual robot is
very hard (specifying link geometry, defining constraints between links, etc).
Therefore we limited ourselves to path planning for a point in $\mathbb{R}^n$,
and we approximated obstacles as being convex polytopes. This is a valid
simplification of the robot motion planning problem for several reasons. The
first reason is that most real robots actually operate in a C-space similar to
$\mathbb{R}^n$. This is because $\mathbb{RP}^3$, the space in which the
rotation of a rigid body occurs, simplifies to $\mathbb{R}^3$ if the joint
cannot make a full rotation. Second, in defining the C-space, the obstacle
space takes into account the interaction of the robot geometry with obstacles.
Therefore motion planning through the free space can be done using a point to
represent the robot (since the geometry of the robot is implicitly included in
the obstacle space). Therefore, although our simulations do not represent real
robots and real obstacles, they do approximate them.

\subsection{Experimental Setup}

We implemented an n-dimensional world in which to test these algorithms. We
structured our code so that we could run our tests in any space of dimension $d
\geq 2$. We also made visualization scripts for the 2 and 3 dimensional tests.
Examples of randomly generated 2D and 3D scenes with candidate solution paths
can be seen in Fig.~\ref{fig:examples}.

Next we made a random obstacle generator. We create n-dimensional convex
polytopes by sampling $n^{2}$ random points within the $n$-sphere. To generate
a single point, you must sample $n - 1$ angles from a uniform distribution
between 0 and $2\pi$, $n$ radii from a Gaussian distribution,
$\mathcal{N}(r_{\mu}, r_{\sigma})$ where $r_{\mu}$ and $r_{\sigma}$ are the
mean and standard deviation for the radius of the convex polytope, and $n$
floats to act as the center of the $n$-sphere.  Then you use the series of
equations in Eq.~\ref{eq:nsphere} to produce a randomly sampled point within
the $n$-sphere where $c_k$, $r_k$, and $\phi_k$ are the centers, radii, and
angles respectively.  To produce a convex polytope, we then computed the
$n$-dimensional convex hull over these points. The convex hull then provided
the normal equations and offsets for each facet on the polytope.

\begin{align}
    \begin{split}
        x_1 &= c_1 + r_1 \cos(\phi_1) \\
        x_2 &= c_2 + r_2 \sin(\phi_1) \cos(\phi_2) \\
        x_3 &= c_3 + r_3 \sin(\phi_1) \sin(\phi_2) \cos(\phi_3) \\
            &\vdots\\
        x_{n-1} &= c_{n - 1} + r_{n - 1} \sin(\phi_1) \cdots \sin(\phi_{n-2}) \cos(\phi_{n-1}) \\
        x_n &= c_n + r_{n} \sin(\phi_1) \cdots \sin(\phi_{n-2}) \sin(\phi_{n-1}) \,.
    \end{split}
    \label{eq:nsphere}
\end{align}

\begin{algorithm}[h!] 
    \caption{$\Function{Dimensionality Reduction}$}
    \label{alg:DR}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}

        \STATE Specify dimension, dimensionality reduction technique, and obstacle information
        \STATE Generate random obstacles based off of obstacle information
        \STATE Fit dimensionality reduction transform to the space
        \STATE Transform the space
        \STATE Run PRM* or RRTConnect over the space
        \STATE Inverse transform the path back to the original dimension
        \STATE Check for collisions
        \RETURN Path\_length, duration, num\_failed
    \end{algorithmic}
\end{algorithm}

After generating random obstacles, we ran Feature Agglomeration, Truncated SVD,
PCA, and Randomized PCA to reduce the dimensionality of the space by using the
sampled points within each $n$-sphere as training data. We then ran PRM* and
RRTConnect on each of these reduced dimensionality spaces, as well as on the
fully-dimensional space as a control. We then ran an inverse transform on the
paths that we find to return them to the fully-dimensional space in order to
verify that the path in the reduced dimension did not come into collision with
obstacles. We collected data on the number of runs that fail to find a
solution, the time it takes to run the algorithm, and the length of the
solution if one was found. Note, that the length of the path is computed after
it has been transformed back into the original high dimensional space in order
to make a fair comparison. A summary of this paragraph can be found in
Algo.~\ref{alg:DR}.

For our experiments, we varied the number of random obstacles in the scene, the
planning algorithm used, and the method of dimensionality reduction. For each
problem instance, the random generator was seeded in order to make fair
comparisons between the planners and dimensionality reduction techniques.
Also, since the algorithms are non-deterministic, each problem instance
executed 100 times. We used scenes with 250, 300, and 350 obstacles in order to
measure how the planners and dimensionality reduction techniques vary in their
performance with varying levels of obstacle density. In the results presented,
we are reducing from 8 dimensions to 6 dimensions and using mean radius of
$14.8$ for the convex polytopes.

\section{Results \& Discussion}

Fig.~\ref{fig:plots} shows plots of various measures of performance using
different methods of dimensionality reduction for motion planing. Please note
that the Y-axes of these plots are not equivalent in the left and right column.
The left and right columns show the results for RRTConnect and PRM algorithms
respectively.  The first, second, and third row show the average path length,
computational time in seconds, and number of failed runs out of 100
respectively. For each dimensionality reduction technique, there are three bars
presenting the results for 250, 300, and 350 randomly generated obstacles
respectively. From the data we can see that both planners produced similarly
sized paths for each dimensionality reduction technique, and that by not
reducing the dimension, the planner is able to find a path roughly two times
shorter.

\begin{figure}[h!]
    \centering
    \makebox[\textwidth][c]{\includegraphics[width=1.5\textwidth]{plots}}

    \caption{Plots showing the length of the solution paths if found, the
    time needed to find the solution, and the number of failed attempts out of
    100 for each method of dimensionality reduction as well as in the high
    dimensional space. The column of plots on the left shows the results for
    the RRTConnect algorithm and the plots on the right show the results for
PRM. Note, The Y-axes limits are not equal and that abbreviations were used to
save space.  F.  Agg.  stands for Feature Agglomeration and R. PCA stands for
Randomized PCA.}

    \label{fig:plots}

\end{figure}

In terms of computational time, there is no significant difference between
Truncated SVD, PCA, Randomized PCA, planning in the original space when
planning with the RRTConnect algorithm. Whereas using Feature Agglomeration
took slightly longer to compute for each set number of obstacles. When using
PRM, the computational time was significantly improved by reducing the
dimension for all methods.  Based on the data, using PRM without dimensionality
reduction took roughly twice as long to compute for each method. This differing
result between PRM and RRTConnect is likely due to the fact that PRM is a
multi-query planning algorithm meaning that it constructs a data-structure that
can be used for multiple planning queries whereas RRTConnect is only looking to
find a single path. This means that the PRM algorithm tries to fully represent
the connectivity of the environment whilst RRTConnect needs to determine the
connectivity directed towards a goal. The reason dimensionality reduction
significantly improves the computational time for PRM is that it is much faster
to exhaust the search space at lower dimensions as opposed to higher
dimensions.

The RRTConnect algorithm had a much lower failure rate than PRM in all
respective dimensionality reduction technique. This is due to the fact that
RRTConnect is growing a tree towards the goal and PRM is randomly sampling and
connecting points to a growing graph. For both algorithms, a similar trend
exists for failure rates when planning in the reduced dimension space. For both
planners, Feature Agglomeration failed more frequently than the rest of the
techniques, and planning in the high dimensional space had the smallest failure
rate. The higher failure rate for Feature Agglomeration is because Feature
Agglomeration works best when features from the data can be combined due to
repetition and redundancy. Since this data was in the $\mathbb{R}^n$ and the
random obstacles were generated all over the space, Feature Agglomeration
failed to produce an embedding that also worked for planning in the lower
dimensional space. Aside from Feature Agglomeration, the dimensionality
reduction techniques failed to find a path to the goal configuration between 20
and 60 percent of the time for PRM and between 5 and 35 percent for RRTConnect.
This is due to the projection of the higher dimensional obstacles into a lower
dimension causing occlusions in the lower dimensional space where they would
not exist in the higher dimension. However, these dimensionality reduction
techniques can be seen as independent random variables for each run, therefore,
you would be able to run the algorithm again with the same probability of
failure. As the number of obstacles increased, so did the failure rate because
the lower dimensional spaces become more densely populated with obstacles
causing more occlusions.

From the data, we can see that dimensionality reduction techniques can reduce
the optimality of the planning algorithm, but in multi-query planning
algorithms, these techniques can significantly reduce the time needed to
compute a solution. Based on the data, it seems that the Truncated SVD
algorithm can be used to produce the lowest number of failures and produces
similar results in the path length and computation time  as compared to the
other techniques used. Also, we Can see that Feature Agglomeration fails to
find a solution path more frequently than the other techniques, but produces
shorter paths when it does find a path.

\section{Conclusion}

\bibliographystyle{plain}
\bibliography{bib}

\end{document}